\vspace{-0.25em}\thetitle

### 2009-07-22 paper.tex
...

\section{Introduction}

@@

\section{Background \& Motivation}

The most fundamental tool in the investigation of linear structure in high dimensions is the singular value decomposition (\caps{SVD}).
This matrix decomposition can be applied to any real matrix and factors it into a product of three matrices:
\begin{align}
  X = US\trans{V}
\end{align}
where $U$ and $V$ are orthogonal and $S$ is diagonal, with entries of decreasing nonnegative value.
Among the useful properties of this factorization, if all but the first $k$ diagonal entries of $S$ are zeroed out, the product $US_k\trans{V}$ is the best rank-$k$ approximation of $X$ under the Frobenius matrix norm.
The rows of $US_k$ can thus be viewed as an optimal $k$-dimensional simplification of the rows of $X$ in the new coordinate system given by the columns of $V$.

\begin{figure}[!b]
\vspace{0.5em}
\begin{center}
\subfloat[IETF 60]{\includegraphics[width=1.6in]{svd/ie60}}
\subfloat[IETF 67]{\includegraphics[width=1.6in]{svd/ie67}}
\subfloat[SIGCOMM 2004]{\includegraphics[width=1.6in]{svd/sc04}}
\subfloat[UCSD]{\includegraphics[width=1.6in]{svd/ucsd}}
\caption{SVD scatter plots for various traces.} 
\fig{svd}
\end{center}
\vspace{-0.85em}
\end{figure}

@@

\section{Methodology}

\section{Results}

\section{Discussion}

\section{Conclusions}

>>>>

\section{Introduction}\sec{introduction}

@@

\section{Background \& Motivation}\sec{background}\sec{motivation}

The most fundamental tool for investigating linear structure in high dimensions is the singular value decomposition (\caps{SVD}).
This matrix decomposition can be applied to any real matrix and factors it into a product of three real matrices:
\begin{align}
  X = US\trans{V}.
\end{align}
Here $U$ and $V$ are orthogonal and $S$ is diagonal, with entries of decreasing nonnegative value.
Among the useful properties of this factorization, if all but the first $k$ diagonal entries of $S$ are zeroed out, the product $US_k\trans{V}$ is the best rank-$k$ approximation of $X$ under the Frobenius matrix norm.
The rows of $US_k$ can thus be viewed as an optimal $k$-dimensional reduced model of the rows of $X$, with a new coordinate system given by the columns of $V$.

We can visualize the $k$-dimensional approximate structure of $X$ in $k$ dimensions using the coordinates of $US_k$.
However, the most significant feature of any vector in this representation is its magnitude.
We are more interested in the direction vectors than their direction:
points with the same direction lie in the same one-dimensional subspace.
To visualize direction while eliminating distance, we project all data points radially onto the hyperplane in which all points have coordinates summing to unity.\footnote{We could normalize the vectors to unit length, but then our projection would be spherical, artificially introducing curvature into our visualization.}
% Then we simply scatter-plot various pairs of dimensions of the resulting data.
\Figure{svd} shows scatter plots of the most significant two projected \caps{SVD} coordinates for four different traces.
The original coordinates of this data encode the packet size distributions, inter-packet interval distributions, \caps{IP} protocol numbers, and port numbers for a sample of 5000 flows for each trace.
The exact details of how the linear representation encodes these features will be discussed in \Section{methodology}.

\begin{figure}[!t]
\vspace{0.5em}
\begin{center}
\subfloat[IETF 60]{\includegraphics[width=1.6in]{svd/ie60}}
\subfloat[IETF 67]{\includegraphics[width=1.6in]{svd/ie67}}
\subfloat[SIGCOMM 2004]{\includegraphics[width=1.6in]{svd/sc04}}
\subfloat[UCSD]{\includegraphics[width=1.6in]{svd/ucsd}}
\caption{SVD scatter plots for various traces.} 
\fig{svd}
\end{center}
\vspace{-0.85em}
\end{figure}

@@

\section{Methodology}\sec{methodology}

\section{Results}\sec{results}

\section{Discussion}\sec{discussion}

\section{Conclusions}\sec{conclusions}
